RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-11-07 discussion.tex
small fixes conclusion

In order to solve these problems, we present RationalGRL: a framework for integrated goal modeling and argumentation. We extend the well-known goal modeling language GRL with argument schemes and critical questions that can be used to analyze and guide stakeholders' discussions about goal models. Our approach, thus, provides a rationalization for the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted while others have been rejected. Our list of argument schemes was constructed by performing an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system. This ensures that the argument schemes we propose are close to actual discussions the stakeholders have (\textbf{requirement 1}).

The meta-model of the RationalGRL framework clearly specifies the traceability links between the arguments based on the schemes and the GRL models (\textbf{requirement 2}). In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}. We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models. These formal traceability links allow us to compute the effect of the arguments and counterarguments proposed in a discussion on a GRL model (\textbf{requirement 3}). In other words, we can determine whether the elements of a GRL model are acceptable given potentially contradictory opinions of stakeholders. Thus, we add a new formal evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (in addition to their \emph{satisfiability}~\cite{Amyot:2010:EGM:1841349.1841356}).

>>>>

In order to solve these problems, we present RationalGRL: a framework for integrated goal modeling and argumentation. We extend the well-known goal modeling language GRL with argument schemes and critical questions that can be used to analyze and guide stakeholders' discussions about goal models. Our approach, thus, provides a rationalization for the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted while others have been rejected. In the introduction, we identified five important requirements for our framework. Below, we discuss how the RationalGRL framework meets these requirements.

Our list of argument schemes and critical questions in Table~\ref{table:argument-schemes} was constructed by performing an extensive case study in which we analyzed 153 pages of transcripts of discussions among designers of a traffic simulator information system. In this way, we ensure that he argumentation techniques capture the actual discussions of the stakeholders or designers in the early requirements engineering phase (\textbf{requirement 1}).

The metamodel of the RationalGRL framework in Figure~\ref{fig:metamodel} clearly specifies the formal traceability links elements of the GRL goal model and the underlying arguments. (\textbf{requirement 2}). In addition to this metamodel, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}. We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models. These formal traceability links allow us to compute the effect of the arguments and counterarguments proposed in a discussion on a GRL model (\textbf{requirement 3}). In other words, we can determine whether the elements of a GRL model are acceptable given potentially contradictory opinions of stakeholders. Thus, we add a new formal evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (in addition to their \emph{satisfiability}~\cite{Amyot:2010:EGM:1841349.1841356}).
